The universal property of infinite direct sums in C$^*$-categories and W$^*$-categories

### 2019-06-06 direct_sums.tex
Add some intro to C* and W*-cats and fix a typo

\begin{thm}[{e.g.~\cite[Proposition~1.2.4]{borceux}}]
\label{directsum_equiv_trad}
Let $\cat{C}$ be an additive category and $A_1,A_2\in\cat{C}$ two objects. Then the following pieces of structure are equivalent for an object $A\in\cat{A}$:
\begin{enumerate}
	\item\label{coprod} $A$ represents the functor
		\[
			\cat{C} \longrightarrow \cat{Ab}, \qquad B \longmapsto \cat{C}(A_1, B) \times \cat{C}(A_2, B)
		\]
	\item\label{prod} $A$ represents the functor
		\[
			\cat{C}^\op \longrightarrow \cat{Ab}, \qquad B \longmapsto \cat{C}(A_1, B) \times \cat{C}(A_2, B)
		\]
	\item\label{biprod} There are morphisms $p_i : A \to A_i$ and $\kappa_i : A_i \to A$ satisfying the \emph{biproduct equations}
	\[
		p_i \kappa_j = \delta_{ij}, \qquad \kappa_1 p_1 + \kappa_2 p_2 = 1.
	\]
\end{enumerate}
\end{thm}

@@

\todo[inline]{State the basic definitions}

>>>>

\begin{thm}[{e.g.~\cite[Proposition~1.2.4]{borceux}}]
\label{directsum_equiv_trad}
Let $\cat{C}$ be an additive category and $A_1,A_2\in\cat{C}$ two objects. Then the following pieces of structure are equivalent for an object $A\in\cat{A}$:
\begin{enumerate}
	\item\label{coprod} $A$ represents the functor
		\[
			\cat{C} \longrightarrow \cat{Ab}, \qquad B \longmapsto \cat{C}(A_1, B) \times \cat{C}(A_2, B)
		\]
    \item\label{prod} $A$ represents the functor
		\[
			\cat{C}^\op \longrightarrow \cat{Ab}, \qquad B \longmapsto \cat{C}(B, A_1) \times \cat{C}(B, A_2)
		\]
	\item\label{biprod} There are morphisms $p_i : A \to A_i$ and $\kappa_i : A_i \to A$ satisfying the \emph{biproduct equations}
	\[
		p_i \kappa_j = \delta_{ij}, \qquad \kappa_1 p_1 + \kappa_2 p_2 = 1.
	\]
\end{enumerate}
\end{thm}

@@

Before we define C$^*$- and~W$^*$-category, we give some context.
A C$^*$-category is a caterogical generalization of a unital C$^*$-algebra
    in the sense that in a~C$^*$-category~$\cat{C}$,
    every homset~$\cat{C}(A,A)$ will be a unital~C$^*$-algebra.
Similarly~$\cat{C}(A,A)$ will be a~W$^*$-algebra
    in a~W$^*$-category $\cat{C}$.
As the definitions of~C$^*$- and~W$^*$-categories
    will implicitly define~C$^*$- and W$^*$-algebras,
    we'll make do with a few short comments.
Every subalgebra of the algebra of complex~$n\times n$-matrices
    is a C$^*$- and~W$^*$-algebra.
More generally, every norm closed subalgebra of~$\B{\H}$,
    the bounded operators on a Hilbert space~$\H$,
    is a C$^*$-algebra and, in fact,
    every C$^*$-algebra is isomorphic to such a norm-closed subalgebra
        of~$\B{\H}$ for some Hilbert space~$\H$.
Every weakly closed subalgebra of~$\B{\H}$ is a W$^*$-algebra
    and every W$^*$-algebra is isomorphic to such a weakly closed subalgebra
Every commutative unital C$^*$-algebra is isomorphic to~$C(X)$ for some
    compact Hausdorff space~$X$
        and every~W$^*$-algebra is isomorphic to~$L^\infty(X)$ for some
            complete finite measure space~$X$.
\begin{defn}
    Let~$\cat{C}$ be a category together with a contravariant endofunctor~$*\colon \cat{C} \to \cat{C}^\op$,
    whose action on arrows we'll denote by~$a\mapsto a^*$.
    The following definitions are due to Ghez, Lima and Roberts \cite{wstarcat}.
\begin{enumerate}
\item
    $\cat{C}$ (together with~$*$) is called a \emph{complex~$*$-category} if the following axioms hold.
\begin{enumerate}
    \item $\cat{C}$ is enriched over complex vector spaces ---
            i.e.~every homset~$\cat{C}(A,B)$ is a complex vector space
                and function composition is bilinear.
    \item The contravariant endofunctor~$*$ is identity-on-objects,
            involutive and conjugate-linear ---
         i.e.~$a^*\colon B\to A$, $a^{**}=a$, $(ab)^*=b^*a^*$, $(\lambda a)^* = \overline{\lambda} a^*$ and
            $(a+c)^* = a^* + c^*$ for arrows~$a, c\colon A \to B$, ~$b\colon C \to A$
                and complex number~$\lambda$.
        (This turns~$\cat{C}(A,A)$ into a~$*$-algebra.)
    \item For any~$a\colon A \to B$
            we have~$a=0$ whenever~$a^*a=0$.
                Furthermore, there is a~$b\colon A \to A$
            with~$a^*a=b^*b$ --- equivalently:
            $a^*a$ is positive in the~$*$-algebra~$\cat{C}(A,A)$.
\end{enumerate}
\item
    $\cat{C}$ (together with~$*$) is called a~\emph{normed $*$-category}
            if it is a complex~$*$-category
            where each~$\cat{C}(A,B)$ is a normed space and~$\| ab \| \leq \|a\|\|b\|$
                for any~$a\colon B \to C$ and~$b\colon A \to B$.
\item
    $\cat{C}$ (together with~$*$) is called a~\emph{C$^*$-category}
            if it is a normed~$*$-category
            where every homset is norm-complete
            and~$\|a \|^2 = \|a^*a\|$.
        (This turns~$\cat{C}(A,A)$ into a unital~C$^*$-algebra.)
\item
    $\cat{C}$ is a~\emph{W$^*$-category}
        if it is a~C$^*$-category
        where every homset has a Banach space predual.
        Equivalently~\cite[Prop.~2.15]{wstarcat}, a C$^*$-category~$\cat{C}$
            is a~W$^*$-category
            if every~$\cat{C}(A,A)$ is a W$^*$-algebra
            and~$\cat{C}(A,B)$ is self-dual
            as a Hilbert~$(A,A)$-module
            with inner product~$\langle a,b\rangle \equiv a^*b$.
\end{enumerate}
\end{defn}

\begin{ex}
The category of Hilbert spaces with bounded linear maps between them
    is a W$^*$-category.
\end{ex}
